# SMM641 Revenue Management and Pricing

## Problem Set 1

Aditya Mohanty

### Question 1

#### a)

```{r FCFS Revenue}
mB=50 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=50 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:1){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
RevenueFCFS=ExpRevenue[1]
print(paste("Lower Bound for Expected Revenue (FCFS) is:", round(RevenueFCFS,2)))

```

#### b)

```{r Optimal Protection for Lunch Demand}
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))

```

#### c)

##### Expected Revenue from Reserve Policy

```{r Expected Revenue from Reserve Policy}
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

##### Percentage Increase from Reserve Policy

```{r Percentage Increase}
PercentImprovement=(OptimalExpRevenue-RevenueFCFS)/RevenueFCFS*100
print(paste("The percentage increase daily revenue from the protection (reserve) policy is:",round(PercentImprovement,2),"%"))
```

#### d)

```{r Expected Revenue vs Protection Plot}
# Plotting Expected Revenue vs Protection Level
xaxis=0:capacity
plot(xaxis,ExpRevenue,pch = 13, cex = 0.5,las=1, xaxt="n",
xlab="Croissants Reserved",ylab="Expected Revenue",cex.lab=1.5, cex.axis=1.5, cex.main=1.5)
xticks <- seq(0, capacity, by=2)
axis(side = 1, at = xticks)
axis(side = 1, at = ProtectBest)
lines(c(ProtectBest,ProtectBest),c(0, max(ExpRevenue)),lty=2)
axis(side = 2, at = round(max(ExpRevenue),2),las=1)
lines(c(0,ProtectBest),c(max(ExpRevenue), max(ExpRevenue)),lty=2)
```

##### If Capacity is increased slightly

```{r Capacity 55}
mB=50 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=55 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

```{r Capacity 70}
mB=50 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=70 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

```{r Capacity 80}
mB=50 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=80 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

##### If capacity is significantly increased

```{r Capacity 150}
mB=50 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=150 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:400){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:400){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

```{r Capacity 250}
mB=50 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=250 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:500){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:500){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

##### If Mean Demand For Breakfast is Increased

```{r Breakfast Demand Increased}
mB=70 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=50 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

##### If Mean Demand For Breakfast is Decreased

```{r Breakfast Demand Decreased}
mB=30 # Mean Demand for Breakfast, Poisson
mL=20 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=50 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

##### If Mean Demand For Lunch is Decreased

```{r Lunch Demand Decreased}
mB=50 # Mean Demand for Breakfast, Poisson
mL=10 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=50 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```

##### If Mean Demand for Lunch is Increased

```{r Lunch Demand Increased}
mB=50 # Mean Demand for Breakfast, Poisson
mL=30 # Mean Demand for Lunch, Poisson
pB=1.00 # Price for Breakfast
pL=1.50 # Price for Lunch
capacity=50 # Number of Croissants received daily
ExpRevenue=rep(0,capacity+1)
for (i in 1:(capacity+1)){
protect=i-1
availforBreakfast=capacity-protect;
ExpRevenue[i]=0;
for(dB in 0:200){
soldBreakfast=min(availforBreakfast,dB)
remainforLunch=capacity-soldBreakfast
for(dL in 0:200){
soldLunch=min(remainforLunch,dL)
RevenueThisIter=pB*soldBreakfast+pL*soldLunch
ExpRevenue[i]=ExpRevenue[i]+
RevenueThisIter*dpois(dB,mB)*dpois(dL,mL)
}
}
}
Protectindexbest = which(ExpRevenue == max(ExpRevenue))
ProtectBest=Protectindexbest-1
OptimalExpRevenue=max(ExpRevenue)
print(paste("The Optimal Protection Level for Lunch Demand is:", ProtectBest))
print(paste("The expected daily revenue from the protection (reserve) policy is:", round(OptimalExpRevenue,2)))
```
