# Question 1

Import packages

```{r}
library(nloptr)
library(stargazer)
library(lattice)
library(dplyr)
```

Import data

```{r}
lonCP <- read.csv('CongestionPricing.csv', header = T)
```

Count rows

```{r}
N=nrow(lonCP)
```

Print top rows to verify

```{r}
head(lonCP)
```

Maximum willingness to pay

```{r}
for (i in 1:N){
lonCP$maxWTP[i]=max(lonCP[i,2:3])
}
```

Display top 10 rows to verify

```{r}
lonCP[1:10,]
```

Find maximum price willing to be paid, as the upper bound

```{r}
maxprice=max(lonCP$maxWTP)
```

### a)

Defining empty array variables we will be introducing

```{r}
demand=rep(NA,maxprice)
revenue=rep(NA,maxprice)
```

Find how many people buy at each price level

```{r}
for (p in 1:maxprice){
demand[p]=sum(lonCP$maxWTP>=p)
revenue[p]=p*demand[p]
}

```

Identifying the Best Price and demand at that level

```{r}
revenueBest=max(revenue)
priceBest=which(revenue == revenueBest)
demandBest=demand[priceBest]
print(paste("If a single price is to be charged across all time slots, the optimal price is:£",priceBest))


```

As this is demand in the representative sample of 345 drivers, finding the total number of drivers at that price is:

```{r}
TotalDrive = (demandBest/345)*192000
```

Total Revenue is:

```{r}
TR1 = TotalDrive * 8

print(paste("The Total revenue is:£",TR1))
```

To calculate average speed, the formula is: Average Speed (in km/h) = 30 -- 0.0625 \* (# of cars in thousands, '000)

```{r}
carsNo = TotalDrive/1000

AvSpd =  30 - (0.0625 * carsNo)

print(paste("The average speed is: ",AvSpd))
```

To calculate Emissions per car, and total emissions, and print the emissions per car:

```{r}
if (AvSpd < 25){
    EmPC= 617.5 - (16.7 * AvSpd)
    TotEmPC = EmPC * TotalDrive
  } else {
    EmPC = 235.0 - (1.4 * AvSpd)
    TotEmPC = EmPC * TotalDrive
  }

print(EmPC)
```

With the optimal price in effect, the total number of emissions will be:

```{r}
print(paste("The Total emissions are (in g/km):",TotEmPC))
```

Total emissions with the optimal price of £8 is 49239404 g/km

### b)

Set base price for non peak timing

```{r}
basePrice=7
```

Defining certain empty variables

```{r}
demandNonPeak<-rep(0,maxprice)
demandPeak<-rep(0,maxprice)
revenue<-rep(0,maxprice)


maxsurplusNonPeak<-rep(0,N)
```

For each client we will obtain their maximum WTP and maximum Surplus among the Nonpeak time slots, and also generate a new column and add this information to our data

```{r}
for (i in 1:N){
        maxsurplusNonPeak[i]=max(lonCP[i,c(3)]-basePrice)
    
    

    lonCP$maxsurplusNonPeak[i]=max(lonCP[i,c(3)]-basePrice)
}

lonCP[1:10,]
```

Let's first compute clients' surpluses for Peak across all possible Peak Price choices

```{r}
surplusPeak<-matrix(0,N,maxprice)

for (p in 1:maxprice){
    for (i in 1:N){
        surplusPeak[i,p]=lonCP[i,2]-p
    }
}

colnames(surplusPeak)=paste0("p=",1:maxprice)
```

Find the demand

```{r}
for (p in 1:maxprice){
  demandNonPeak[p]=sum((maxsurplusNonPeak>surplusPeak[,p])*(maxsurplusNonPeak>=0))
  demandPeak[p]=sum((surplusPeak[,p]>=maxsurplusNonPeak)*(surplusPeak[,p]>=0))
  revenue[p]=basePrice*demandNonPeak[p]+p*demandPeak[p]
}
```

Plotting NonPeak Demand vs Peak Period Price

```{r}
xaxis=1:maxprice
plot(xaxis,demandNonPeak,pch = 16, type="s",col="blue", las=1, xaxt="n",
     xlab="Price for Peak Period (Time slot 7am-6pm)",ylab="Non-Peak Period Demand")
xticks <- seq(0, maxprice, by=1)
axis(side = 1, at = xticks)

```

Plotting Peak Demand vs Peak Period Price

```{r}
xaxis=1:maxprice
plot(xaxis,demandPeak,pch = 16, type="s",col="blue", las=1, xaxt="n",
     xlab="Price for Peak Period",ylab="Peak Period Demand")
xticks <- seq(0, maxprice, by=1)
axis(side = 1, at = xticks)
```

Plotting Revenue vs Peak Period Price

```{r}
xaxis=1:maxprice
plot(xaxis,revenue/1000,pch = 16, type="s",col="blue", las=1, xaxt="n",
     xlab="Price for Peak Period",ylab="Total Revenue")
xticks <- seq(0, maxprice, by=1)
axis(side = 1, at = xticks)
revenueBest=max(revenue[basePrice:maxprice])
priceBest=which(revenue == revenueBest)
axis(side = 1, at = priceBest) 
lines(c(priceBest,priceBest),c(0, revenueBest/1000),lty=2)
axis(side = 2, at = round(revenueBest/1000,3),las=1)
lines(c(0,priceBest),c(revenueBest/1000, revenueBest/1000),lty=2)
```

```{r}
print(paste("When other periods have a base price of 7, the optimal price for the peak time slot is:",priceBest))

```

Calculate demand at the Non-peak and peak times

```{r}
BestDemandNonPeak = demandNonPeak[basePrice]
BestDemandPeak = demandPeak[priceBest]
```

Calculate total drivers at peak and non peak times

```{r}
TotalPeak = (BestDemandPeak/345) * 192000
TotalNonPeak = (BestDemandNonPeak/345) * 192000
```

Total Revenue in this scenario is:

```{r}
TR2 = (TotalPeak * 9) + (TotalNonPeak * 7)

print(paste("The Total revenue is:£",TR2))
```

Average speed in this scenario:

```{r}
AvSpd2 = 30 - (0.0625*(TotalPeak + TotalNonPeak)/1000)

print(paste("The Average Speed is:",AvSpd2))
```

Average emissions per car:

```{r}
if (AvSpd2 < 25){
    EmPC2= 617.5 - (16.7 * AvSpd2)
    TotEm2 = EmPC2 * (TotalPeak + TotalNonPeak)
  } else {
    EmPC2 = 235.0 - (1.4 * AvSpd2)
    TotEm2 = EmPC2 * (TotalPeak + TotalNonPeak)
  }

print(paste("The emissions per car are (in g/km):",EmPC2))
```

Total emissions in the scenario:

```{r}
TotEm2 = EmPC2 * (TotalPeak + TotalNonPeak)

print(paste("The Total emissions are (in g/km):",TotEm2))
```

### c)

Set baseprice

```{r}
basePrice=7
```

Defining certain empty variables

```{r}
demandNonPeak<-rep(0,maxprice)
demandPeak<-rep(0,maxprice)
revenue<-rep(0,maxprice)
emissions <- rep(0,maxprice)
maxsurplusNonPeak<-rep(0,N)
TotalPeak2 <-rep(0, maxprice)
TotalNonPeak2 <-rep(0, maxprice)
AvgSpd3 <-rep(0, maxprice)
EPC <-rep(0, maxprice)
TotEPC <-rep(0, maxprice)
TotalCars <-rep(0, maxprice)
Price <-rep(0, maxprice)

```

Obtain max WTP per client

```{r}
for (i in 1:N){
        maxsurplusNonPeak[i]=max(lonCP[i,c(3)]-basePrice)
    

    lonCP$maxsurplusNonPeak[i]=max(lonCP[i,c(3)]-basePrice)
}

lonCP[1:10,]
```

Compute client surplus

```{r}
surplusPeak<-matrix(0,N,maxprice)

for (p in 1:maxprice){
    for (i in 1:N){
        surplusPeak[i,p]=lonCP[i,2]-p
    }
}

colnames(surplusPeak)=paste0("p=",1:maxprice)
```

Compute demand, total number of cars, average speed, emissions per car and total emissions

```{r}
for (p in 1:maxprice){
  demandNonPeak[p]=sum((maxsurplusNonPeak>surplusPeak[,p])*(maxsurplusNonPeak>=0))
  demandPeak[p]=sum((surplusPeak[,p]>=maxsurplusNonPeak)*(surplusPeak[,p]>=0))
  revenue[p]=basePrice*(192000*demandNonPeak[p]/N)+p*(192000*demandPeak[p]/N)
  TotalPeak2[p] = (demandPeak[p]/345) * 192000
  TotalNonPeak2[p] = (demandNonPeak[p]/345) * 192000
  TotalCars[p] = TotalPeak2[p] + TotalNonPeak2[p]
  Price[p]=p
  AvgSpd3[p] = 30 - (0.0625*(TotalPeak2[p]+TotalNonPeak2[p])/1000 )
  if (AvgSpd3[p] < 25){
    EPC[p] = 617.5 - (16.7 * AvgSpd3[p])
    TotEPC[p] = EPC[p] * TotalCars[p]
  } else {
    EPC[p] = 235.0 - (1.4 * AvgSpd3[p])
    TotEPC[p] = EPC[p] * TotalCars[p]
  }
  
}
```

Place all the values in a dataframe to make the data wrangling simpler

```{r}
PriceData <- data.frame(Price,revenue, demandPeak, demandNonPeak,TotalPeak2, TotalNonPeak2, AvgSpd3,EPC,TotEPC)
```

Transfer the values which have acceptable revenue of over £1,100,000 to a new dataframe.

```{r}
AcceptableRev <- PriceData[(PriceData[,2]>1100000),]
```

Find the ideal price which leads to the minimum total emissions.

```{r}
IdealPrice <- AcceptableRev[which.min(AcceptableRev$TotEPC),]
```

Print the resulting optimal price, revenue and total emissions

```{r}
print(paste("The optimal price is:£",IdealPrice$Price))
print(paste("The revenue is:£",IdealPrice$revenue))
print(paste("The Total emissions are (in g/km):",IdealPrice$TotEPC))
```
